Elastic Scattering in General Relativistic Ray Tracing for Neutrinos

### 2017-10-31 paper.tex
Add a few clarifying edits

\begin{abstract}
  We present a covariant ray tracing algorithm for computing high-resolution
  neutrino distributions in general numerical spacetimes with hydrodynamical
  sources.
  Our formulation treats the very important effect of
  elastic scattering of neutrinos off of nuclei and nucleons
  (changing the neutrino's direction but not energy)
  by incorporating estimates of the background neutrino fields.
  These background fields may be taken from a low-order moment simulation
  or be ignored, in which case the method
  reduces to a standard state-of-the-art ray tracing formulation.
  It handles radiation in regimes spanning optically thick to optically thin.
  We test the new code, highlight its strengths and weaknesses,
  and apply it to a simulation of a neutron star merger
  to compute neutrino fluxes and spectra,
  and to demonstrate a neutrino flavor oscillation calculation.
\end{abstract}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{fig-affine_parameter_label}
  % Script: $DIR/figures-radiation_transport_schematics.nb
  % w/ DIR=/Users/mbdeaton/Dropbox/Research/NuclearAstrophysics/NotesAndReference
  \caption{Affine parameterization of a neutrino trajectory of momentum
    $p_\alpha$.
    The fiducial observer with velocity $u^\alpha$ sits at $x^\alpha_o$,
    the neutrino emission event is at $x^\alpha_e$. The affine parameter
    increases from the emission event: $\lambda_e<\lambda'<\lambda''$.}
  \label{fig:affine_param}
\end{figure}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{figure-mfps-50km}
  % Data:   $DIR/table-50km-nu?.dat
  % Script: $DIR/GnuplotScripts/plot-mfp-50km.gnu
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/HomogeneousIsotropic
  \caption{Mean free paths
    representative of the collapse profile at
    50~km ($\varrho=10^{11}\,{\rm g}\,{\rm cm}^{-3}$,
    $T=3.7$~MeV, $Y_e=0.12$).
    In the fluid rest frame these are given by the inverses of the opacities,
    $\chi^*_a=\mathscr{K}_{\rm AE}/\varepsilon$ for absorption and
    $\chi_s=\mathscr{K}_{\rm SE}/\varepsilon$ for elastic scattering interactions;
    see App.~\ref{sec:source_terms} for definitions.
    Note that elastic scattering opacities are identical across species
    below energies at which weak magnetism plays a role \cite{horo2002-weak_mag}.
    In this data they are exactly identical
    because we have turned off weak magnetism in our opacity calculations
    in order to compare our results with the historical literature.
    Computed using \lstinline{NuLib}.
  }
  \label{fig:mfps-50km}
\end{figure}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{fig-cumulative_f-homogeneous}
  % Data:   $DIR/Lev1-v0-ShootANeutrino-01/Lev1
  % Script: $DIR/GnuplotScripts/plot_f-integral_over_t.gnu
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/HomogeneousIsotropic
  \caption{Cumulative distribution functions at $\varepsilon=11.25$~MeV
    in homogeneous infinite slab of Sec.~\ref{ssec:test_equilibrium}.
    The integration proceeds from left to right, or backwards in time $t$.
    Each ray terminates when it achieves a total optical depth
    greater than $\tau_{\rm term}=14$.
    The positions of the points (only plotted for the $f_{\rm AE}$ components)
    are the time steps chosen by the adaptive time-stepping algorithm
    described in Sec.~\ref{ssec:numerical}.
  }
  \label{fig:cumulative_f_homogeneous}
\end{figure}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{figure-ferr-50km}
  % Data:   $DIR/Lev1-v0-noscat-01/Lev1
  % Script: $DIR/GnuplotScripts/plot_ferr.gnu
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/HomogeneousIsotropic
  \caption{Relative error in integrated equilibrium distribution functions
    in the infinite homogeneous slab  described in
    Sec.~\ref{ssec:test_equilibrium}.
    Presented here are the relative differences between the final
    $f_{\rm AE}$ in the \emph{noscat} case
    (see Fig.~\ref{fig:cumulative_f_homogeneous})
    and the equilibrium distribution functions given by
    Eqn.~\ref{eqn:feq}.
    The source of this error is the discarded boundary term
    $f_{\rm bdry}\sim10^{-6}$, described in Sec.~\ref{ssec:numerical}.
  }
  \label{fig:homogeneous_isotropic}
\end{figure}

@@

\begin{figure}
  %\includegraphics[width=\columnwidth]{fig-moving_fluid_avg_eps}
  \resizebox{\columnwidth}{!}{\input{fig-moving_fluid_avg_eps.tex}}
  % Data:   $DIR/Lev1-v*-noscat-01
  % Script: $DIR/plot-avg_eps_vs_cosA.gnu
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/MovingFluid
  \caption{Average energy of $\bar{\nu}_e$ neutrino fields
    measured by a stationary observer,
    $\langle\tilde{\varepsilon}\rangle(\cos A)$.
    The observer sees neutrinos with $\cos A>0$ to be moving primarily with
    the fluid.
    The points are computed from ray tracing spectra; the lines are the
    analytic formula Eqn.~\ref{eqn:doppler_avg_eps}.
  }
  \label{fig:avg_eps_doppler}
\end{figure}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{fig-f-absorption_star-11MeV}
  % Data:   $DIR/Lev1-01/Lev1
  % Script: $DIR/Lev1-01/Lev1/plot_f_monochromes-11.25MeV.gnu
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/HomogeneousAbsorptionSphere
  \caption{Distribution functions outside an idealized homogeneous star
    with radius $R=50$~km, and observer at $r=75$~km,
    described in Sec.~\ref{ssec:test_ab_star}.
    $A$ is the angle between the neutrino momentum and the $\hat{r}$ direction.
    In this plot we display only the samples at energy
    $\varepsilon=11.25\,{\rm MeV}$.
    The points are computed from ray tracing; the lines from the analytic
    solution, Eqn.~\ref{eqn:fae_homogeneous_general}.
  }
  \label{fig:f_absorption_sphere}
\end{figure}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{fig-collapse_profile_avg_eps_vs_cosA}
  % Data:   $DIR/
  % Script: $DIR_GRAY_SC/../plot-compare_spectral_vs_gray-angular_distribution_avg_eps-production.gnu
  % w/  DIR_MC     =$DIR/ResultsSherwood-03/Spectra
  %     DIR_SPEC_SC=$DIR/ResultsBrett-A-03/Test-scat-01
  %     DIR_GRAY_SC=$DIR/ResultsBrett-B-03/Test-scat-01
  %     DIR_GRAY_AB=$DIR/ResultsBrett-B-03/Test-noscat-01
  % and DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/CollapseProfile/
  \caption{Angular distribution of $\nu_x$ average energies
    $\langle\varepsilon\rangle(\cos A)$
    in the collapse profile test described in Sec.~\ref{ssec:test_collapse}.
    The observer is at 500~km, and the shock at 150~km,
    so that the half-opening angle of the shock is $\cos A\approx0.954$.
    The four methods depicted are 1) a fiducial Monte Carlo calculation,
    ray tracing using the 2) spectral and 3) gray methods to estimate background
    fields for scattering, and 4) ray tracing neglecting scattering.
  }
  \label{fig:avg_eps_collapse}
\end{figure}

@@

\begin{table}%[h]
  % to get mc luminosities: tail $MCDATA/Spectra/luminosities_and_avg_eps-501km-nu?.dat
  % to get rt luminosities: tail $RTDATA/Moments-nu?.out
  % MCDATA=$DIR/ResultsSherwood-02
  % RTDATA_NO=$DIR/ResultsBrett-B-03/Test-noscat-01/
  % RTDATA_GR=$DIR/ResultsBrett-B-03/Test-scat-01/
  % RTDATA_SP=$DIR/ResultsBrett-A-03/Test-scat-01/
  % DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/CollapseProfile/
  \caption{
    Comparison of total luminosities and average energies between
    the following methods:
    ray tracing ignoring elastic scattering `RT noscat',
    ray tracing with the gray scattering treatment `RT scat gr',
    ray tracing with the spectral scattering treatment `RT scat sp', and
    a fiducial Monte Carlo transport evolution `MC'.
    We also show average energies of the scattering envelope
    $\langle\tilde{\varepsilon}\rangle^{\rm sc}$, which are estimated by eye
    from plots like Fig.~\ref{fig:avg_eps_collapse} at $\cos A < 0.95$.
    Luminosities have units $10^{52}\,{\rm erg}\,{\rm s}^{-1}$ and
    average energies have units MeV.
    The $\nu_x$ luminosities are per-species:
    multiply by four to get the total heavy-lepton neutrino luminosities.
  }
  \label{tab:collapse_rt_vs_mc}
  \begin{tabularx}{\columnwidth}{X r r r r}
    & {\bf RT noscat} & \,\,{\bf RT scat gr} & \,\,{\bf RT scat sp} & \,\,{\bf MC} \\
    \hline
    $L_{\nu_e}$                                         &  3.53 & 3.14 & 3.04 & 3.48 \\
    $L_{\bar{\nu}_e}$                                   & 5.19 & 3.05 & 3.03 & 3.01 \\
    $L_{\nu_x}$                                         & 222.0  & 1.88 & 1.76 & 1.70 \\
    $\langle \tilde{\varepsilon}_{\nu_e} \rangle$               &  10.6 & 10.6 & 10.9 & 11.0 \\
    $\langle \tilde{\varepsilon}_{\bar{\nu}_e} \rangle$         & 14.2 & 13.0 & 13.8 & 13.7 \\
    $\langle \tilde{\varepsilon}_{\nu_x} \rangle$               & 47.8 & 16.0 & 17.3 & 16.2 \\
    $\langle \tilde{\varepsilon}_{\nu_e} \rangle^{\rm sc}$      & 4    & 14   & 17   & 16 \\
    $\langle \tilde{\varepsilon}_{\bar{\nu}_e} \rangle^{\rm sc}$& 4    & 24   & 20   & 18 \\
    $\langle \tilde{\varepsilon}_{\nu_x} \rangle^{\rm sc}$      & 2    & 28   & 31   & 27 \\
    \hline
  \end{tabularx}
\end{table}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{production-colormap-merid-rho}
  % VTK Data: zosima:/home/mbdeaton/Research/20170602-nsns_vol_data
  % Plots:    $DIR/production-colormap-{merid,equat}-{rho,temp}.png
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170420-nsns_volume_data/02-two_domains_avg_eps/PlotsEtc
  \caption{A meridional slice of density in the hypermassive neutron
    star and disk configuration.
    The distorted rectangular boundaries are the boundaries of the grid
    used in the numerical simulation, which employs a coordinate mapping to
    concentrate points near the central object.}
  \label{fig:nsns_rho_merid}
\end{figure}

\begin{figure}
  \includegraphics[width=\columnwidth]{production-colormap-equat-rho}
  % see notes from production-colormap-merid-rho
  \caption{An equatorial slice of density in the hypermassive neutron
    star and disk configuration.}
  \label{fig:nsns_rho_equat}
\end{figure}

\begin{figure}
  \includegraphics[width=\columnwidth]{production-colormap-merid-temp}
  % see notes from production-colormap-merid-rho
  \caption{A meridional slice of temperature in the hypermassive neutron
    star and disk configuration.}
  \label{fig:nsns_temp_merid}
\end{figure}

\begin{figure}
  \includegraphics[width=\columnwidth]{production-colormap-equat-temp}
  % see notes from production-colormap-merid-rho
  \caption{An equatorial slice of temperature in the hypermassive neutron
    star and disk configuration.}
  \label{fig:nsns_temp_equat}
\end{figure}

@@

\begin{table}%[h]
  % Total fluxes over the sky integrated in google sheet: 20170726-theta_distribution_integrations
  % RTDATA_NO=$DIR/PositionDistributions-nu?-250km-Lev1-e30a120b20-noscat-01.dat
  % RTDATA_GR=$DIR/PositionDistributions-nu?-250km-Lev1-e30a60b10-scat-01.dat
  % DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170324-ray_tracing_nsns/Run03/PositionDistributions
  \caption{
    Comparison of total luminosities and average energies
    of the hypermassive neutron star configuration.
    The methods are ray tracing using the \emph{noscat} method `RT noscat',
    ray tracing using the \emph{gray} \emph{scat} method `RT scat',
    and the M1 transport simulation `M1'.
    The ray tracing totals were computed from sums over
    observers placed in the $y$-$z$ plane.
    M1 values are taken from \cite[Figs. 7, 9, 10]{fouc2016-m1_evolve_n}.
    Energy luminosities have units $10^{52}\,{\rm erg}\,{\rm s}^{-1}$,
    number luminosities $10^{57}\,{\rm s}^{-1}$, and
    average energies MeV.
    The $\nu_x$ luminosities are per-species:
    multiply by four to get the total heavy-lepton neutrino luminosities.
  }
  \label{tab:nsns_rt_vs_m1}
  \begin{tabularx}{\columnwidth}{X r r r}
    & {\bf RT noscat} & \,\,{\bf RT scat} & \,\,{\bf M1} \\
    \hline
    $L_{\nu_e}$                                 & 5.62 & 6.05 & 5 \\
    $L_{\bar{\nu}_e}$                           & 10.6 & 11.8 & 12 \\
    $L_{\nu_x}$                                 & 23.3 & 12.7 & 12 \\
    $R_{\nu_e}-R_{\bar{\nu}_e}$                 & -1.22 & -1.85 & -2 \\
    $\langle \varepsilon_{\nu_e} \rangle$       & 12.08 & 11.91 & 12 \\
    $\langle \varepsilon_{\bar{\nu}_e} \rangle$ & 16.07 & 14.67 & 15 \\
    $\langle \varepsilon_{\nu_x} \rangle$       & 34.09 & 23.58 & 26 \\
    \hline
  \end{tabularx}
\end{table}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{theta_distrib-250km-Kr}
  % Data_a:   $DIR/PositionDistributions/*Lev1-e30a120b20-01*
  % Data_s:   $DIR/PositionDistributions/*Lev1-e30a60b10-01*
  % Data gathered using: $DIR/CollectPositionDistributions.sh
  % Script: $DIR/GnuScripts/plot-theta_distributions-Hr-fixed_r.gnu
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170324-ray_tracing_nsns/Run03
  \caption{Radial number fluxes, $K^r(\theta)$,
    as defined in Eqn.~\ref{eqn:Kr_minkowski},
    in the hypermassive neutron star configuration.
    Sampled for observers at fixed coordinate distances from the center
    of the star, with $r=250\,{\rm km}$.
  }
  \label{fig:nsns_theta_distrib_Kr}
\end{figure}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{cosA_distrib-intensity-250km-00deg}
  % Data:   $DIR/Lev1-e30a60b10-{,no}scat-00
  % Script: $DIR/plot-angular_distribution-intensity-compare.gnu
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170324-ray_tracing_nsns/Run03/000Mo-000.0deg-250km
  \caption{Angular distributions of fluxes for an observer on the rotation axis
    in the hypermassive neutron star configuration.}
  \label{fig:nsns_cosA_distrib_I_00deg}
\end{figure}

\begin{figure}
  \includegraphics[width=\columnwidth]{cosA_distrib-intensity-250km-90deg}
  % Data:   $DIR/Lev1-e30a60b10-{,no}scat-00
  % Script: $DIR/plot-angular_distribution-intensity-compare.gnu
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170324-ray_tracing_nsns/Run03/000Mo-090.0deg-250km
  \caption{Angular distributions of fluxes for an observer in the equatorial
    plane in the hypermassive neutron star configuration.}
  \label{fig:nsns_cosA_distrib_I_90deg}
\end{figure}

>>>>

\begin{abstract}
  We present a covariant ray tracing algorithm for computing high-resolution
  neutrino distributions in general relativistic numerical spacetimes with
  hydrodynamical sources.
  Our formulation treats the very important effect of
  elastic scattering of neutrinos off of nuclei and nucleons
  (changing the neutrino's direction but not energy)
  by incorporating estimates of the background neutrino fields.
  These background fields may be taken from a low-order moment simulation
  or be ignored, in which case the method
  reduces to a standard state-of-the-art ray tracing formulation.
  It handles radiation in regimes spanning optically thick to optically thin.
  We test the new code, highlight its strengths and weaknesses,
  and apply it to a simulation of a neutron star merger
  to compute neutrino fluxes and spectra,
  and to demonstrate a neutrino flavor oscillation calculation.
\end{abstract}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{fig-affine_parameter_label}
  % Script: $DIR/figures-radiation_transport_schematics.nb
  % w/ DIR=/Users/mbdeaton/Dropbox/Research/NuclearAstrophysics/NotesAndReference
  \caption{Affine parameterization of a neutrino trajectory of momentum
    $p_\alpha$.
    The fiducial observer with velocity $u^\alpha$ sits at $x^\alpha_o$,
    the neutrino emission event is at $x^\alpha_e$. The affine parameter
    increases from the emission event: $\lambda_e<\lambda'<\lambda''$.
    The green slab represents dense matter.}
  \label{fig:affine_param}
\end{figure}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{figure-mfps-50km}
  % Data:   $DIR/table-50km-nu?.dat
  % Script: $DIR/GnuplotScripts/plot-mfp-50km.gnu
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/HomogeneousIsotropic
  \caption{Mean free paths
    representative of the post-bounce collapse profile (presented in
    Sec.~\ref{ssec:test_collapse}) at
    50~km, where $\varrho=10^{11}\,{\rm g}\,{\rm cm}^{-3}$,
    $T=3.7$~MeV, and $Y_e=0.12$; this thermodynamic state is used in
    the tests with homogeneous matter distributions
    (presented in Secs.~\ref{ssec:test_equilibrium}--\ref{ssec:test_ab_star}).
    In the fluid rest frame the mean free paths are given by the inverses of the opacities
    $\chi^*_a=\mathscr{K}_{\rm AE}/\varepsilon$ for absorption and
    $\chi_s=\mathscr{K}_{\rm SE}/\varepsilon$ for elastic scattering interactions
    (see App.~\ref{sec:source_terms} for definitions).
    Note that elastic scattering opacities are identical across species
    below energies at which weak magnetism plays a role \cite{horo2002-weak_mag}.
    In this data they are exactly identical
    because we have turned off weak magnetism in our opacity calculations
    in order to compare our results with the historical literature.
    Computed using \lstinline{NuLib}.
  }
  \label{fig:mfps-50km}
\end{figure}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{fig-cumulative_f-homogeneous}
  % Data:   $DIR/Lev1-v0-ShootANeutrino-01/Lev1
  % Script: $DIR/GnuplotScripts/plot_f-integral_over_t.gnu
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/HomogeneousIsotropic
  \caption{Cumulative distribution functions at $\varepsilon=11.25$~MeV
    in the homogeneous infinite slab test
    (presented in Sec.~\ref{ssec:test_equilibrium}).
    The integration proceeds from left to right, or backwards in time $t$.
    Each ray terminates when it achieves a total optical depth
    greater than $\tau_{\rm term}=14$.
    The positions of the points (only plotted for the $f_{\rm AE}$ components)
    are the time steps chosen by the adaptive time-stepping algorithm
    described in Sec.~\ref{ssec:numerical}.
  }
  \label{fig:cumulative_f_homogeneous}
\end{figure}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{figure-ferr-50km}
  % Data:   $DIR/Lev1-v0-noscat-01/Lev1
  % Script: $DIR/GnuplotScripts/plot_ferr.gnu
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/HomogeneousIsotropic
  \caption{Relative error in integrated equilibrium distribution functions
    in the infinite homogeneous slab
    (presented in Sec.~\ref{ssec:test_equilibrium}).
    Plotted here are the relative differences between the final
    $f_{\rm AE}$ in the \emph{noscat} case
    (see Fig.~\ref{fig:cumulative_f_homogeneous})
    and the equilibrium distribution functions given by
    Eqn.~\ref{eqn:feq}.
    The source of this error is the discarded boundary term
    $f_{\rm bdry}\sim10^{-6}$, described in Sec.~\ref{ssec:numerical}.
  }
  \label{fig:homogeneous_isotropic}
\end{figure}

@@

\begin{figure}
  %\includegraphics[width=\columnwidth]{fig-moving_fluid_avg_eps}
  \resizebox{\columnwidth}{!}{\input{fig-moving_fluid_avg_eps.tex}}
  % Data:   $DIR/Lev1-v*-noscat-01
  % Script: $DIR/plot-avg_eps_vs_cosA.gnu
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/MovingFluid
  \caption{Average energy of $\bar{\nu}_e$ neutrino fields
    measured by a stationary observer,
    $\langle\tilde{\varepsilon}\rangle(\cos A)$,
    in the moving slab test (presented in Sec.~\ref{ssec:test_doppler}).
    The observer sees neutrinos with $\cos A>0$ to be moving primarily with
    the fluid.
    The points are computed from ray tracing spectra; the lines are the
    analytic formula Eqn.~\ref{eqn:doppler_avg_eps}.
  }
  \label{fig:avg_eps_doppler}
\end{figure}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{fig-f-absorption_star-11MeV}
  % Data:   $DIR/Lev1-01/Lev1
  % Script: $DIR/Lev1-01/Lev1/plot_f_monochromes-11.25MeV.gnu
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/HomogeneousAbsorptionSphere
  \caption{Distribution functions outside an idealized homogeneous star
    with radius $R=50$~km, and observer at $r=75$~km
    (presented in Sec.~\ref{ssec:test_ab_star}).
    $A$ is the angle between the neutrino momentum and the $\hat{r}$ direction.
    In this plot we display only the samples at energy
    $\varepsilon=11.25\,{\rm MeV}$.
    The points are computed from ray tracing; the lines from the analytic
    solution, Eqn.~\ref{eqn:fae_homogeneous_general}.
  }
  \label{fig:f_absorption_sphere}
\end{figure}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{fig-collapse_profile_avg_eps_vs_cosA}
  % Data:   $DIR/
  % Script: $DIR_GRAY_SC/../plot-compare_spectral_vs_gray-angular_distribution_avg_eps-production.gnu
  % w/  DIR_MC     =$DIR/ResultsSherwood-03/Spectra
  %     DIR_SPEC_SC=$DIR/ResultsBrett-A-03/Test-scat-01
  %     DIR_GRAY_SC=$DIR/ResultsBrett-B-03/Test-scat-01
  %     DIR_GRAY_AB=$DIR/ResultsBrett-B-03/Test-noscat-01
  % and DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/CollapseProfile/
  \caption{Angular distribution of $\nu_x$ average energies
    $\langle\varepsilon\rangle(\cos A)$
    in the collapse profile test (described in Sec.~\ref{ssec:test_collapse}).
    The observer is at 500~km, and the shock at 150~km,
    so that the half-opening angle of the shock is $\cos A\approx0.954$.
    The four methods depicted are 1) a fiducial Monte Carlo calculation,
    ray tracing using the 2) spectral and 3) gray methods to estimate background
    fields for scattering, and 4) ray tracing neglecting scattering.
  }
  \label{fig:avg_eps_collapse}
\end{figure}

@@

\begin{table}%[h]
  % to get mc luminosities: tail $MCDATA/Spectra/luminosities_and_avg_eps-501km-nu?.dat
  % to get rt luminosities: tail $RTDATA/Moments-nu?.out
  % MCDATA=$DIR/ResultsSherwood-02
  % RTDATA_NO=$DIR/ResultsBrett-B-03/Test-noscat-01/
  % RTDATA_GR=$DIR/ResultsBrett-B-03/Test-scat-01/
  % RTDATA_SP=$DIR/ResultsBrett-A-03/Test-scat-01/
  % DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/CollapseProfile/
  \caption{
    From the post-bounce collapse profile test (Sec.~\ref{ssec:test_collapse}),
    a comparison of total luminosities and average energies between
    the following methods:
    ray tracing ignoring elastic scattering `RT noscat',
    ray tracing with the gray scattering treatment `RT scat gr',
    ray tracing with the spectral scattering treatment `RT scat sp', and
    a fiducial Monte Carlo transport evolution `MC'.
    We also show average energies of the scattering envelope
    $\langle\tilde{\varepsilon}\rangle^{\rm sc}$, which are estimated by eye
    from plots like Fig.~\ref{fig:avg_eps_collapse} at $\cos A < 0.95$.
    Luminosities have units $10^{52}\,{\rm erg}\,{\rm s}^{-1}$ and
    average energies have units MeV.
    The $\nu_x$ luminosities are per-species:
    multiply by four to get the total heavy-lepton neutrino luminosities.
  }
  \label{tab:collapse_rt_vs_mc}
  \begin{tabularx}{\columnwidth}{X r r r r}
    & {\bf RT noscat} & \,\,{\bf RT scat gr} & \,\,{\bf RT scat sp} & \,\,{\bf MC} \\
    \hline
    $L_{\nu_e}$                                         &  3.53 & 3.14 & 3.04 & 3.48 \\
    $L_{\bar{\nu}_e}$                                   & 5.19 & 3.05 & 3.03 & 3.01 \\
    $L_{\nu_x}$                                         & 222.0  & 1.88 & 1.76 & 1.70 \\
    $\langle \tilde{\varepsilon}_{\nu_e} \rangle$               &  10.6 & 10.6 & 10.9 & 11.0 \\
    $\langle \tilde{\varepsilon}_{\bar{\nu}_e} \rangle$         & 14.2 & 13.0 & 13.8 & 13.7 \\
    $\langle \tilde{\varepsilon}_{\nu_x} \rangle$               & 47.8 & 16.0 & 17.3 & 16.2 \\
    $\langle \tilde{\varepsilon}_{\nu_e} \rangle^{\rm sc}$      & 4    & 14   & 17   & 16 \\
    $\langle \tilde{\varepsilon}_{\bar{\nu}_e} \rangle^{\rm sc}$& 4    & 24   & 20   & 18 \\
    $\langle \tilde{\varepsilon}_{\nu_x} \rangle^{\rm sc}$      & 2    & 28   & 31   & 27 \\
    \hline
  \end{tabularx}
\end{table}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{production-colormap-merid-rho}
  % VTK Data: zosima:/home/mbdeaton/Research/20170602-nsns_vol_data
  % Plots:    $DIR/production-colormap-{merid,equat}-{rho,temp}.png
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170420-nsns_volume_data/02-two_domains_avg_eps/PlotsEtc
  \caption{A meridional slice of density in the hypermassive neutron
    star and disk configuration (Sec.~\ref{ssec:test_disk_comparison}).
    The distorted rectangular boundaries are the boundaries of the grid
    used in the numerical simulation, which employs a coordinate mapping to
    concentrate points near the central object.}
  \label{fig:nsns_rho_merid}
\end{figure}

\begin{figure}
  \includegraphics[width=\columnwidth]{production-colormap-equat-rho}
  % see notes from production-colormap-merid-rho
  \caption{An equatorial slice of density in the hypermassive neutron
    star and disk configuration (Sec.~\ref{ssec:test_disk_comparison}).}
  \label{fig:nsns_rho_equat}
\end{figure}

\begin{figure}
  \includegraphics[width=\columnwidth]{production-colormap-merid-temp}
  % see notes from production-colormap-merid-rho
  \caption{A meridional slice of temperature in the hypermassive neutron
    star and disk configuration (Sec.~\ref{ssec:test_disk_comparison}).}
  \label{fig:nsns_temp_merid}
\end{figure}

\begin{figure}
  \includegraphics[width=\columnwidth]{production-colormap-equat-temp}
  % see notes from production-colormap-merid-rho
  \caption{An equatorial slice of temperature in the hypermassive neutron
    star and disk configuration (Sec.~\ref{ssec:test_disk_comparison}).}
  \label{fig:nsns_temp_equat}
\end{figure}

@@

\begin{table}%[h]
  % Total fluxes over the sky integrated in google sheet: 20170726-theta_distribution_integrations
  % RTDATA_NO=$DIR/PositionDistributions-nu?-250km-Lev1-e30a120b20-noscat-01.dat
  % RTDATA_GR=$DIR/PositionDistributions-nu?-250km-Lev1-e30a60b10-scat-01.dat
  % DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170324-ray_tracing_nsns/Run03/PositionDistributions
  \caption{
    Comparison of total luminosities and average energies
    of the hypermassive neutron star configuration
    (presented in Sec.~\ref{ssec:test_disk_comparison}).
    The methods are ray tracing using the \emph{noscat} method `RT noscat',
    ray tracing using the \emph{gray} \emph{scat} method `RT scat',
    and the M1 transport simulation `M1'.
    The ray tracing totals were computed from sums over
    observers placed in the $y$-$z$ plane.
    M1 values are taken from \cite[Figs. 7, 9, 10]{fouc2016-m1_evolve_n}.
    Energy luminosities have units $10^{52}\,{\rm erg}\,{\rm s}^{-1}$,
    number luminosities $10^{57}\,{\rm s}^{-1}$, and
    average energies MeV.
    The $\nu_x$ luminosities are per-species:
    multiply by four to get the total heavy-lepton neutrino luminosities.
  }
  \label{tab:nsns_rt_vs_m1}
  \begin{tabularx}{\columnwidth}{X r r r}
    & {\bf RT noscat} & \,\,{\bf RT scat} & \,\,{\bf M1} \\
    \hline
    $L_{\nu_e}$                                 & 5.62 & 6.05 & 5 \\
    $L_{\bar{\nu}_e}$                           & 10.6 & 11.8 & 12 \\
    $L_{\nu_x}$                                 & 23.3 & 12.7 & 12 \\
    $R_{\nu_e}-R_{\bar{\nu}_e}$                 & -1.22 & -1.85 & -2 \\
    $\langle \varepsilon_{\nu_e} \rangle$       & 12.08 & 11.91 & 12 \\
    $\langle \varepsilon_{\bar{\nu}_e} \rangle$ & 16.07 & 14.67 & 15 \\
    $\langle \varepsilon_{\nu_x} \rangle$       & 34.09 & 23.58 & 26 \\
    \hline
  \end{tabularx}
\end{table}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{theta_distrib-250km-Kr}
  % Data_a:   $DIR/PositionDistributions/*Lev1-e30a120b20-01*
  % Data_s:   $DIR/PositionDistributions/*Lev1-e30a60b10-01*
  % Data gathered using: $DIR/CollectPositionDistributions.sh
  % Script: $DIR/GnuScripts/plot-theta_distributions-Kr-fixed_r.gnu
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170324-ray_tracing_nsns/Run03
  \caption{Radial number fluxes, $K^r(\theta)$,
    as defined in Eqn.~\ref{eqn:Kr_minkowski},
    in the hypermassive neutron star configuration
    (presented in Sec.~\ref{ssec:test_disk_comparison}).
    Sampled for observers at fixed coordinate distances from the center
    of the star, with $r=250\,{\rm km}$.
  }
  \label{fig:nsns_theta_distrib_Kr}
\end{figure}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{cosA_distrib-intensity-250km-00deg}
  % Data:   $DIR/Lev1-e30a60b10-{,no}scat-00
  % Script: $DIR/plot-angular_distribution-intensity-compare.gnu
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170324-ray_tracing_nsns/Run03/000Mo-000.0deg-250km
  \caption{Angular distributions of fluxes for an observer on the rotation axis
    in the hypermassive neutron star configuration
    (presented in Sec.~\ref{ssec:test_disk_comparison}).}
  \label{fig:nsns_cosA_distrib_I_00deg}
\end{figure}

\begin{figure}
  \includegraphics[width=\columnwidth]{cosA_distrib-intensity-250km-90deg}
  % Data:   $DIR/Lev1-e30a60b10-{,no}scat-00
  % Script: $DIR/plot-angular_distribution-intensity-compare.gnu
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170324-ray_tracing_nsns/Run03/000Mo-090.0deg-250km
  \caption{Angular distributions of fluxes for an observer in the equatorial
    plane in the hypermassive neutron star configuration
    (presented in Sec.~\ref{ssec:test_disk_comparison}).}
  \label{fig:nsns_cosA_distrib_I_90deg}
\end{figure}
